Structured Versus Decorated Cospans

### 2021-06-21 cospans_new.tex
completed proof of Thm. 3.1.

Lastly, we show that $\Psi \maps L/\X \to \A$ satisfies the Beck--Chevalley condition. 
For this, given a pushout square in $\A$:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1,0) {$b$};
\node (C) at (0,-1) {$c$};
\node (D) at (1,-1) {$d$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$h$} (B)
(A) edge node[left]{$k$} (C)
(B) edge node[right]{$g$} (D)
(C) edge node[above]{$f$} (D);
\end{tikzpicture}
\]
we need to show that the natural transformation 
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$(L/\X)_a$};
\node (B) at (1.5,0) {$(L/\X)_b$};
\node (C) at (0,-1) {$(L/\X)_c$};
\node (D) at (1.5,-1) {$(L/\X)_d$};
\node (E) at (0.75,-0.5) {$\Searrow \theta$};
\path[->,font=\scriptsize,>=angle 90]
(B) edge node[above]{$h^*$} (A)
(A) edge node[left]{$k_!$} (C)
(B) edge node[right]{$g_!$} (D)
(D) edge node[below]{$f^*$} (C);
\end{tikzpicture}
\]
defined in Equation \eqref{theta} is an isomorphism.  Let $(b,L(b) \xrightarrow{i} x)$ be an object of $(L/\X)_b$.  Going left, we precompose with $L(h) \maps L(a) \to L(b)$ to obtain the object $(a,L(a) \xrightarrow{i  L(h)} x)$ in $(L/\X)_a$.  Then going down, we push forward along $L(k) \maps L(a) \to L(c)$:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$L(a)$};
\node (B) at (1.5,0) {$x$};
\node (C) at (0,-1) {$L(c)$};
\node (D) at (1.5,-1) {$L(c)+_{L(a)}x$};
\node at (1.1,-0.7) {\Large{$\ulcorner$}};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$I \circ L(h)$} (B)
(A) edge node[left]{$L(k)$} (C)
(B) edge node[right]{$$} (D)
(C) edge node[above]{$\psi$} (D);
\end{tikzpicture}
\]
to obtain the object $(c,L(c) \xrightarrow{\psi} L(c)+_{L(a)} x)$ of $(L/ \X)_c$. For the other route, first going down, we push the object $(b,x,L(b) \xrightarrow{I} x)$ forward along $L(g) \maps L(b) \to L(d)$ by taking the following pushout:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$L(b)$};
\node (B) at (1.5,0) {$x$};
\node (C) at (0,-1) {$L(d)$};
\node (D) at (1.5,-1) {$L(d)+_{L(b)}x$};
\node at (1.1,-0.7) {\Large{$\ulcorner$}};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$I$} (B)
(A) edge node[left]{$L(g)$} (C)
(B) edge node[right]{$$} (D)
(C) edge node[above]{$\phi$} (D);
\end{tikzpicture}
\]
which yields the object $(d,L(d) \xrightarrow{\theta} L(d)+_{L(b)} x)$ of $(L/\X)_d$.  Then going left, we precompose with $L(f) \maps L(c) \to L(d)$ to obtain the object $(c,L(c) \xrightarrow{\phi L(f)} L(d)+_{L(b)} x)$ of $(L/\X)_c$.

To establish a natural isomorphism between the two objects $(c,L(c)+_{L(a)} x,L(c) \xrightarrow{\psi} L(c)+_{L(a)} x)$ and $(c,L(d)+_{L(b)} x,L(c) \xrightarrow{\theta \circ L(f)} L(d)+_{L(b)} x)$ in $(L/\X)_c$, it suffices to establish a natural isomorphism between the objects $L(c)+_{L(a)} x$ and $L(d)+_{L(b)} x$ in $\X$. The first pushout square above exhibits $L(c)+_{L(a)}x$ as a colimit of the span $$L(c) \xleftarrow{L(k)} L(a) \xrightarrow{I \circ L(h)} x$$in $\X$. On the other hand, as $L$ preserves finite colimits, $L$ preserves pushout squares, and so the following square is also a pushout:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$L(a)$};
\node (B) at (1,0) {$L(b)$};
\node (C) at (0,-1) {$L(c)$};
\node (D) at (1,-1) {$L(d)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$L(h)$} (B)
(A) edge node[left]{$L(k)$} (C)
(B) edge node[right]{$L(g)$} (D)
(C) edge node[above]{$L(f)$} (D);
\end{tikzpicture}
\]
which we may then paste to the left of the second above pushout square to obtain another pushout square:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$L(b)$};
\node (B) at (2,0) {$x$};
\node (C) at (0,-1) {$L(d)$};
\node (D) at (2,-1) {$L(d)+_{L(b)}x$};
\node (E) at (-1,0) {$L(a)$};
\node (F) at (-1,-1) {$L(c)$};
\path[->,font=\scriptsize,>=angle 90]
(E) edge node [left] {$L(k)$} (F)
(E) edge node [above] {$L(h)$} (A)
(F) edge node [above] {$L(f)$} (C)
(A) edge node[above]{$I$} (B)
(A) edge node[left]{$L(g)$} (C)
(B) edge node[right]{$$} (D)
(C) edge node[above]{$\theta$} (D);
\end{tikzpicture}
\]
which exhibits $L(d)+_{L(b)} x$ also as a colimit of the same span in $\X$, and hence the desired natural isomorphism between our two objects of $(L /\X)_c$. Thus $\Psi$ satisfies the Beck--Chevalley condition.
\end{proof}

>>>>

Lastly, we show that $\Psi \maps L/\X \to \A$ satisfies the Beck--Chevalley condition. 
For this, given a pushout square in $\A$:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1,0) {$b$};
\node (C) at (0,-1) {$c$};
\node (D) at (1,-1) {$d$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$h$} (B)
(A) edge node[left]{$k$} (C)
(B) edge node[right]{$g$} (D)
(C) edge node[below]{$f$} (D);
\end{tikzpicture}
\]
we need to show that the natural transformation 
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$(L/\X)_a$};
\node (B) at (1.5,0) {$(L/\X)_b$};
\node (C) at (0,-1) {$(L/\X)_c$};
\node (D) at (1.5,-1) {$(L/\X)_d$};
\node (E) at (0.75,-0.5) {$\Searrow \theta$};
\path[->,font=\scriptsize,>=angle 90]
(B) edge node[above]{$h^*$} (A)
(A) edge node[left]{$k_!$} (C)
(B) edge node[right]{$g_!$} (D)
(D) edge node[below]{$f^*$} (C);
\end{tikzpicture}
\]
defined in Equation \eqref{theta} is an isomorphism.  Let $(b,L(b) \xrightarrow{i} x)$ be an object of $(L/\X)_b$.  Going left, we precompose with $L(h) \maps L(a) \to L(b)$ to obtain the object $(a,L(a) \xrightarrow{i  L(h)} x)$ in $(L/\X)_a$.  Then going down, we push forward along $L(k) \maps L(a) \to L(c)$:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$L(a)$};
\node (B) at (1.5,0) {$x$};
\node (C) at (0,-1) {$L(c)$};
\node (D) at (1.5,-1) {$L(c)+_{L(a)}x$};
\node at (1.1,-0.7) {\Large{$\ulcorner$}};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$I \circ L(h)$} (B)
(A) edge node[left]{$L(k)$} (C)
(B) edge node[right]{$$} (D)
(C) edge node[below]{$\psi$} (D);
\end{tikzpicture}
\]
to obtain the object $(c,L(c) \xrightarrow{\psi} L(c)+_{L(a)} x)$ of $(L/ \X)_c$. For the other route, first going down, we push the object $(b,x,L(b) \xrightarrow{I} x)$ forward along $L(g) \maps L(b) \to L(d)$ by taking the following pushout:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$L(b)$};
\node (B) at (1.5,0) {$x$};
\node (C) at (0,-1) {$L(d)$};
\node (D) at (1.5,-1) {$L(d)+_{L(b)}x$};
\node at (1.1,-0.7) {\Large{$\ulcorner$}};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$I$} (B)
(A) edge node[left]{$L(g)$} (C)
(B) edge node[right]{$$} (D)
(C) edge node[below]{$\phi$} (D);
\end{tikzpicture}
\]
which yields the object $(d,L(d) \xrightarrow{\theta} L(d)+_{L(b)} x)$ of $(L/\X)_d$.  Then going left, we precompose with $L(f) \maps L(c) \to L(d)$ to obtain the object $(c,L(c) \xrightarrow{\phi L(f)} L(d)+_{L(b)} x)$ of $(L/\X)_c$.

The natural transformation $\theta$ gives a morphism
\[   \theta \maps (c,L(c) \xrightarrow{\psi} L(c)+_{L(a)} x) \; \to \; (c,L(c) \xrightarrow{\phi  L(f)} L(d)+_{L(b)} x) \]
in $(L/\X)_c$.   As in the proof of \cref{thm:decorated_cospans_1}, this arises from the universal property of the pushout:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$L(b)$};
\node (B) at (1.5,0) {$x$};
\node (C) at (0,-1) {$L(d)$};
\node (D) at (1.5,-1) {$L(d)+_{L(b)}x$};
\node (E) at (-1.5,0) {$L(a)$};
\node (F) at (-1.5,-1) {$L(c)$};
\node (G) at (2.5,-2) {$L(c)+_{L(a)}x$};
\node at (1.1,-0.7) {\Large{$\ulcorner$}};
\path[->,font=\scriptsize,>=angle 90]
(E) edge node [left] {$L(k)$} (F)
(E) edge node [above] {$L(h)$} (A)
(F) edge node [below] {$L(f)$} (C)
(A) edge node[above]{$i$} (B)
(A) edge node[left]{$L(g)$} (C)
(B) edge node[right]{$$} (D)
(C) edge node[below]{$\phi$} (D)
(F) edge [bend right] node[below]{$\psi$} (G)
(G) edge [dashed] node[below left]{$\exists ! \;\zeta $} (D)
(B) edge [bend left] (G);
\end{tikzpicture}
\]
Namely, we have $\theta = (1,\zeta)$.   But when the original square in $A$ is a pushout,
the square at left above is also a pushout, since $L$ preserves finite colimits.  Since the pasting
of pushout squares is a pushout, $\zeta$ and hence $\theta$ is an isomorphism, so $\Psi$ satisfies the Beck--Chevalley condition.
\end{proof}
